Reproducible Research Peer Assignment 1
=============================================

###Loading and Preprocessing the data

1. Loading the activity.csv file.
```{r echo = T}
activity <- read.csv ("E:/Documents/Data Science Spl/Reproducible Research/Peer Assg 1/activity.csv")
head (activity)
```

2. Pre-Processing the data (Changing  class of date variable)
```{r}
activity$date <- as.Date (activity$date)
```

###What is mean total number of steps taken per day?

1. Removing all NA values
```{r}
act_good <- activity [ complete.cases(activity),]
head (act_good , n=  3)
```

2. Calculating total number of steps each day
```{r}
total <- aggregate (steps~date, act_good, sum)
head (total , n =3)
```

3. Plotting a histogram and printing mean and median values
```{r , fig.height= 4}
hist (total$steps ,breaks = 10, col = "red" , xlab = "Total Number of Step"  , main = "Total Steps in a Day")
mean (total$steps)
median (total$steps)
```

###What is the average daily activity pattern?
1. Aggregating by interval and Calculating Mean
```{r}
avg_int <- aggregate (steps~interval, act_good , mean)
head (avg_int)
```

2. Plotting the mean
```{r}
with (avg_int , plot (steps~interval , type = "l" , xlab = "Interval" , ylab = "Average Steps"))
```

3. Checking which 5 minute interval has maximum average steps
```{r}
index <- which.max (avg_int$steps)
avg_int$interval[index]
```

###Imputing missing values

1. Getting the number of missing values
```{r}
idx <- is.na (activity$step)
sum (idx)
```

We will use the mean number of steps per 5 minute interval to fill out the missing values.

2. Copying datasets into another data frame and  Replacing those NA values with  average values corresponding to same intervals

```{r , cache=TRUE}
activity_mod <- activity
for (i in idx)
    {
    activity_mod$steps[idx] = avg_int$steps[ avg_int$interval == activity_mod$interval[idx]]
    }
head (activity_mod)
```

3. Plotting histogram and calculating mean and median of total number of steps per day

```{r ,fig.height= 4}
total2 <- aggregate (steps~date , activity_mod , sum)
head (total2)
hist (total2$steps ,breaks = 10, col = "red" , xlab = "Total Number of Step"  , main = "Total Steps in a Day")
mean (total2$steps)
median (total2$steps)
```

Compared to original data , mean has not changed but median has seen a very small change.

###Are there differences in activity patterns between weekdays and weekends?

1. Determining the type of day
```{r , cache=TRUE}
activity_mod$wd <- weekdays (activity_mod$date)
for (i in 1:nrow(activity_mod))
    {
    if (activity_mod$wd[i] == "Sunday" | activity_mod$wd[i] == "Saturday")
        activity_mod$wd[i] = "weekend"
    else 
        activity_mod$wd[i] = "weekday"
}
head (activity_mod$wd)
```

2. Converting it to a factor
```{r}
activity_mod$wd <- as.factor (activity_mod$wd)
head (activity_mod$wd)
```

3. Plotting the Graph
```{r, fig.height= 6}
library(ggplot2)
avg_int_wd <- aggregate (steps~interval+wd,activity_mod , mean)
head (avg_int_wd)
ggplot (avg_int_wd , aes (x = interval  , y = steps)) +geom_line() + facet_grid (wd~.)+ xlab ("Interval") + ylab ("Average Number of Steps") + theme_bw()+ ggtitle("Weekend and Weekday Distribution")
```

As clearly seen from the plot, the average activity on weekend is lower than on weekdays.


